### 2019-05-08 inst/response_to_reviewers.md

# Response to reviewers

- We have changed the installation method to use `remotes` (which has lower
  dependencies than `devtools`).
- We have added disk location to `cites_status`
- We now clear out and disconnect the database tables after updating, which
  should trigger disk cleanup and avoid doubling database size

- We have opted not to use **glue** but stick with `paste` in the interest of
  limiting dependencies.  This is a trade-off but we believe a minor one.
- The low test coverage shown in https://github.com/ecohealthalliance/citesdb/issues/4
  is due to tests skippeed on CRAN.  Setting the environment variable to 
  `NOT_CRAN=true` shows that our test coverage is 65%.  This is lower than
  typical, but as we note above, this is largely due to the verbose code
  for interacting with the RStudio connection pane, which can not be
  straightforwardly testeed. Other code coverage is 
  [greater than 90%](https://codecov.io/gh/ecohealthalliance/citesdb/tree/master/R)
  and all major funcionality is tested.

### 2019-05-08 inst/response_to_reviewers.md

# Response to reviewers

Thanks to both reviewers for their comments and issuees

## Changes implemented

- We have changed the installation method to use `remotes` (which has lower
  dependencies than `devtools`).
- We have added disk location to `cites_status`
- We now clear out and disconnect the database tables after updating, which
  should trigger disk cleanup and avoid doubling database size
- We have modified out linter tests to avoid the false positives shown in
  https://github.com/ecohealthalliance/citesdb/issues/2
- We have renamed `cites_status()` to `cites_check_status()` to be less generic

## Changes not implemented / justifications

- We have opted not to use **glue** but stick with `paste` in the interest of
  limiting dependencies.  This is a trade-off but we believe a minor one.
- The low test coverage shown in https://github.com/ecohealthalliance/citesdb/issues/4
  is due to tests skippeed on CRAN.  Setting the environment variable to 
  `NOT_CRAN=true` shows that our test coverage is 65%.  This is lower than
  typical, but as we note above, this is largely due to the verbose code
  for interacting with the RStudio connection pane, which can not be
  straightforwardly tested. Other code coverage is 
  [greater than 90%](https://codecov.io/gh/ecohealthalliance/citesdb/tree/master/R)
  and all major funcionality is tested.

### 2019-05-08 inst/response_to_reviewers.md

# Response to reviewers

Thanks to both reviewers for their comments and issues

## Changes implemented

- We have changed the installation method to use `remotes` (which has lower
  dependencies than `devtools`).
- We have added disk location to `cites_status`
- We now clear out and disconnect the database tables after updating, which
  should trigger disk cleanup and avoid doubling database size
- We have modified out linter tests to avoid the false positives shown in
  https://github.com/ecohealthalliance/citesdb/issues/2
- We have renamed `cites_status()` to `cites_check_status()` to be less generic
- We have made more elaborate help and examples for the `cites_db()`, 
  `cites_shipments()`, and metadata functions to illustrate their use and
  distinguish between **dplyr** and **DBI**-based workflows.

@@

## Changes not implemented / justifications

- We have opted not to use **glue** but stick with `paste` in the interest of
  limiting dependencies.  This is a trade-off but we believe a minor one.
- The low test coverage shown in https://github.com/ecohealthalliance/citesdb/issues/4
  is due to tests skippeed on CRAN.  Setting the environment variable to 
  `NOT_CRAN=true` shows that our test coverage is 65%.  This is lower than
  typical, but as we note above, this is largely due to the verbose code
  for interacting with the RStudio connection pane, which can not be
  straightforwardly tested. Other code coverage is 
  [greater than 90%](https://codecov.io/gh/ecohealthalliance/citesdb/tree/master/R)
  and all major funcionality is tested.

We have not yet been able to recreate the error in
https://github.com/ecohealthalliance/citesdb/issues/4, but have made some
changes that may resolve it.  Can reviewers test this and report back in that
issue?

A note - we have heard from the maintainers of MonetDBLite that it will not be
returning to CRAN, but they are working on a successor embeedddeed database package that
will replace it and go to CRAN later this year. So for now we will host this
package on GitHub and replace the database back-end and send to CRAN when
the successor is ready.

### 2019-05-08 inst/response_to_reviewers.md

## Changes not implemented / justifications

- We have opted not to use **glue** but stick with `paste` in the interest of
  limiting dependencies.  This is a trade-off but we believe a minor one.
- The low test coverage shown in https://github.com/ecohealthalliance/citesdb/issues/4
  is due to tests skippeed on CRAN.  Setting the environment variable to 
  `NOT_CRAN=true` shows that our test coverage is 65%.  This is lower than
  typical, but as we note above, this is largely due to the verbose code
  for interacting with the RStudio connection pane, which can not be
  straightforwardly tested. Other code coverage is 
  [greater than 90%](https://codecov.io/gh/ecohealthalliance/citesdb/tree/master/R)
  and all major funcionality is tested.

We have not yet been able to recreate the error in
https://github.com/ecohealthalliance/citesdb/issues/4, but have made some
changes that may resolve it.  Can reviewers test this and report back in that
issue?

A note - we have heard from the maintainers of MonetDBLite that it will not be
returning to CRAN (it's current iteration fails on R-devel), but they are
working on a successor embeedddeed database package that will replace it and go
to CRAN later this year. So for now we will host this package on GitHub and
replace the database back-end and send to CRAN when the successor is ready.

### 2019-05-08 inst/response_to_reviewers.md

## Changes implemented

- We have changed the installation method to use `remotes` (which has lower
  dependencies than `devtools`).
- We have added disk location to `cites_status`
- We now clear out and disconnect the database tables after updating, which
  should trigger disk cleanup and avoid doubling database size
- We have modified out linter tests to avoid the false positives shown in
  https://github.com/ecohealthalliance/citesdb/issues/2
- We have renamed `cites_status()` to `cites_check_status()` to be less generic
- We have made more elaborate help and examples for the `cites_db()`, 
  `cites_shipments()`, and metadata functions to illustrate their use and
  distinguish between **dplyr** and **DBI**-based workflows.

## Changes not implemented / justifications

- We have opted not to use **glue** but stick with `paste` in the interest of
  limiting dependencies.  This is a trade-off but we believe a minor one.
- The low test coverage shown in https://github.com/ecohealthalliance/citesdb/issues/4
  is due to tests skippeed on CRAN.  Setting the environment variable to 
  `NOT_CRAN=true` shows that our test coverage is 65%.  This is lower than
  typical, but as we note above, this is largely due to the verbose code
  for interacting with the RStudio connection pane, which can not be
  tested other than in an interactive session. Other code coverage is 
  [greater than 90%](https://codecov.io/gh/ecohealthalliance/citesdb/tree/master/R)
  and all major funcionality is tested.

### 2019-05-08 inst/response_to_reviewers.md

## Changes not implemented / justifications

- We have opted not to use **glue** but stick with `paste` in the interest of
  limiting dependencies.  This is a trade-off but we believe a minor one.
- The low test coverage shown in https://github.com/ecohealthalliance/citesdb/issues/4
  is due to tests skippeed on CRAN.  Setting the environment variable to 
  `NOT_CRAN=true` shows that our test coverage is 65%.  This is lower than
  typical, but as we note above, this is largely due to the verbose code
  for interacting with the RStudio connection pane, which can not be
  tested other than in an interactive session. Other code coverage is 
  [greater than 90%](https://codecov.io/gh/ecohealthalliance/citesdb/tree/master/R)
  and all major funcionality is tested.

We have not yet been able to recreate the error in
https://github.com/ecohealthalliance/citesdb/issues/4, but have made some
changes that may resolve it.  Can reviewers test this and report back in that
issue?

A note - we have heard from the maintainers of MonetDBLite that it will not be
returning to CRAN (it's current iteration fails on R-devel), but they are
working on a successor embedded database package that will replace it and go
to CRAN later this year. So for now we will host this package on GitHub and
replace the database back-end and send to CRAN when the successor is ready.

### 2019-05-08 inst/response_to_reviewers.md

## Changes implemented

- We have changed the installation method to use `remotes` (which has lower
  dependencies than `devtools`).
- We have added disk location to `cites_status`
- We now clear out and disconnect the database tables after updating, which
  should trigger disk cleanup and avoid doubling database size
- We have modified out linter tests to avoid the false positives shown in
  https://github.com/ecohealthalliance/citesdb/issues/2
- We have renamed `cites_status()` to `cites_check_status()` to be less generic
- We have made more elaborate help and examples for the `cites_db()`, 
  `cites_shipments()`, and metadata functions to illustrate their use and
  distinguish between **dplyr** and **DBI**-based workflows.

## Changes not implemented / justifications

- We have opted not to use **glue** but stick with `paste` in the interest of
  limiting dependencies.  This is a trade-off but we believe a minor one.
- The low test coverage shown in https://github.com/ecohealthalliance/citesdb/issues/4
  is due to tests skippeed on CRAN.  Setting the environment variable to 
  `NOT_CRAN=true` shows that our test coverage is 65%.  This is lower than
  typical, but as we note above, this is largely due to the verbose code
  for interacting with the RStudio connection pane, which can not be
  tested other than in an interactive session. Other code coverage is 
  [greater than 90%](https://codecov.io/gh/ecohealthalliance/citesdb/tree/master/R).
  We believe that all important functionality is tested, including important
  edge-cases and conditions not reflected in the coverage statistic, such as
  error handling in multiple sessions and changing up upstream data sources.

### 2019-05-08 inst/response_to_reviewers.md

# Response to reviewers

Thanks to both reviewers for their comments and issues that have 
helped us to make a more user-friendly package.

@@

## Changes implemented

- We have changed the installation method in the `README` to use **remotes**, 
  which has fewer dependencies than **devtools**

- We have added information to the `cites_status` table indicating the location 
  of the database files on disk

- We now clear out and disconnect the database tables after updating, which
  should trigger disk cleanup and avoid doubling database size

- We have modified our linter tests to avoid the false positives shown in
  https://github.com/ecohealthalliance/citesdb/issues/2

- We have renamed the internal function `check_status()` to `cites_check_status()`
  to be less generic

- We have made more elaborate help and examples for the `cites_db()`, 
  `cites_shipments()`, and metadata functions to illustrate their use and
  distinguish between **dplyr** and **DBI**-based workflows.

@@

## Changes not implemented / justifications

- We have opted not to use the **glue** package and instead stick with the base
  R `paste()` functions in the interest of limiting dependencies. We believe
  this is a minor trade-off.

- The low test coverage shown in https://github.com/ecohealthalliance/citesdb/issues/4
  is due to tests skipped on CRAN. Setting the environment variable to 
  `NOT_CRAN=true` shows that our test coverage is 65%. This is lower than
  typical, but as we note in the issue above, this is largely due to the 
  verbose code for interacting with the RStudio connection pane, which cannot be
  tested except in an interactive session. Other code coverage is 
  [greater than 90%](https://codecov.io/gh/ecohealthalliance/citesdb/tree/master/R).
  We believe that all important functionality is tested, including important
  edge-cases and conditions not reflected in the coverage statistic, such as
  error handling in multiple sessions and changing up upstream data sources.

We have not yet been able to recreate the error in
https://github.com/ecohealthalliance/citesdb/issues/1, but have made some
changes that may resolve it. Can reviewers test this and report back in that
issue?

A final note: we have heard from the maintainers of **MonetDBLite** that it will 
not be returning to CRAN (its current iteration fails on R-devel), but they are
working on a successor embedded database package that will replace it and go
to CRAN later this year. So, for now, we will host this package on GitHub and
replace the database back-end and send to CRAN when the successor package is ready.

### 2019-05-08 README.Rmd
Various edits to review fixes

When you first load the package you will see a message like this:

@@

Once you fetch the data you can connect to the database with the `cites_db()` command. You can use the `cites_shipments()` command to load a remote `tibble` that is backed by the database but not loaded into R. You can use this to analyze CITES data without ever loading it into memory, then gather your results with the `dplyr` function `collect()`. For example:

@@

The back-end database, [MonetDB](https://monetdb.org), is very fast and powerful, making such analyses quite snappy even on such large data using normal desktops and laptops. Here's the timing of the above query, which processes over 20 million records:

@@

If you don't need any of the bells and whistles of this package, you can download the raw data as a single compressed TSV file from the [releases page](https://github.com/ecohealthalliance/citesdb/releases), or as a `.zip` file of many CSV files from original source at <https://trade.cites.org/>.

@@

```{r message = FALSE, warning = FALSE, error = FALSE, include = FALSE}
citesdb::cites_disconnect()
MonetDBLite::monetdblite_shutdown()
```

>>>>

When you first load the package, you will see a message like this:

@@

Once you fetch the data, you can connect to the database with the `cites_db()` command. The `cites_shipments()` command loads a remote `tibble` that is backed by the database but is not loaded into R. You can use this command to analyze CITES data without ever loading it into memory, gathering your results with the `dplyr` function `collect()`. For example:

@@

The back-end database, [MonetDB](https://monetdb.org), is very fast and powerful, making analyses on such large data quite snappy using normal desktops and laptops. Here's the timing of the above query, which processes over 20 million records:

@@

If you don't need any of the bells and whistles of this package, you can download the raw data as a single compressed TSV file from the [releases page](https://github.com/ecohealthalliance/citesdb/releases), or as a `.zip` file of many CSV files from the original source at <https://trade.cites.org/>.

@@

```{r message = FALSE, warning = FALSE, error = FALSE, include = FALSE}
citesdb::cites_disconnect()
```
